---
title: "Data Science Homework 3"
author: "Yujin Zhang"
date: "10/17/2021"
output: github_document
---

```{r}
library(tidyverse)
library(httr)
library(jsonlite)

knitr::opts_chunk$set(
  warning = FALSE,
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)
```

## This is a R Markdown document for Data Science Homework 3.

Problem 1

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.
```{r}
# load the dataset
library(p8105.datasets)
data("instacart")

head(instacart, 4)
```
Description:

The dataset includes 15 variables and 1384617 observations. The above table shows the example rows of this dataset, where each row in the dataset is a product from an order. The meaning of every variable is:

order_id: order identifier

product_id: product identifier

add_to_cart_order: order in which each product was added to cart

reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise

user_id: customer identifier

eval_set: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from the “train” eval_set)

order_number: the order sequence number for this user (1=first, n=nth)

order_dow: the day of the week on which the order was placed

order_hour_of_day: the hour of the day on which the order was placed

days_since_prior_order: days since the last order, capped at 30, NA if order_number=1

product_name: name of the product

aisle_id: aisle identifier

department_id: department identifier

aisle: the name of the aisle

department: the name of the department


1. How many aisles are there, and which aisles are the most items ordered from?
```{r}
# count the number of aisles
count_aisle = 
  instacart %>% 
  group_by(aisle) %>%
  count(name = "n_obs") %>% 
  arrange(desc(n_obs))
  
head(count_aisle, 2)
```

For this question, the amount of aisles was 134, and "fresh vegetables" was the most items ordered from.

2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. 
```{r}
# filter the aisle whose count > 10000
count_filter = 
  instacart %>% 
  group_by(aisle_id, aisle) %>% 
  count(name = "n_obs") %>% 
  filter(n_obs > 10000) %>% 
  arrange(desc(n_obs))

# make a plot to show the number of items ordered in each aisle  
ggplot(count_filter, aes(x = reorder(aisle, n_obs), y = n_obs)) +
  geom_bar(aes(fill = aisle), stat = "identity") +
  labs(title = "bar diagram for number of items ordered in each aisle", x = "name of aisles", y = "count") +
  scale_x_discrete(labels = NULL)
```

The bar diagram above showed the number of items ordered in aisles, which had more than 10000 items ordered. 

3. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
# make a table showing the three most popular
join_most3popular =
  instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits" )) %>% 
  group_by(aisle, product_name) %>% 
  count(name = "counts") %>% 
  group_by(aisle) %>% 
  mutate(rank =  order(order(counts, decreasing = TRUE))) %>% # add a variable which shows rank of each product
  filter(rank < 4) %>% # find the three most popular items
  arrange(aisle)
  

```
`r knitr::kable(join_most3popular)`

4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
```{r}
# select variables and observations in Pink Lady Apples and Coffee Ice Cream
order_hour = 
  instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>%
  arrange(product_name, order_dow) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  mutate(order_dow = recode(order_dow, "0" = "Sunday",
                            "1" = "Monday",
                            "2" = "Tuesday",
                            "3" = "Wednesday",
                            "4" = "Thursday",
                            "5" = "Friday",
                            "6" = "Saturday")) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  )  # format a readable table 
  
```
`r knitr::kable(order_hour)`

## Problem 2

Load the dataset
```{r}
var_names = 
  GET("https://chronicdata.cdc.gov/views/acme-vg9e.json") %>%
  content("text") %>%
  fromJSON() %>% 
  .[["columns"]] %>% 
  .[["name"]] %>% 
  .[1:22]

brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/views/acme-vg9e/rows.json") %>% 
  content("text") %>%
  fromJSON() %>% 
  .[["data"]]

row_as_tibble = function(row_as_list, var_names) {
  var_list = row_as_list[9:30]
  names(var_list) = var_names 
  var_list[sapply(var_list, is.null)] <- NULL
  as_tibble(var_list, validate = FALSE)
}

brfss_smart2010 = 
  brfss_smart2010 %>% 
  map(.x = ., ~row_as_tibble(.x, var_names)) %>% 
  bind_rows
```

1. do data cleaning:

Format the data to use appropriate variable names;
focus on the “Overall Health” topic;
include only responses from “Excellent” to “Poor”;
organize responses as a factor taking levels ordered from “Poor” to “Excellent”;

```{r}
data_brfss_smart2010 =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename("location_abbr" = locationabbr, "location_desc" = locationdesc, "resp_id" = respid) %>% 
  separate(location_desc, into = c("state", "location"), sep = " - ") %>% 
  filter(topic == "Overall Health") %>% 
  filter(response %in% c("Excellent", "Poor", "Very good", "Good", "Fair")) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  mutate(data_value = as.numeric(data_value)) %>% 
  arrange(year) %>% 
  select(-location_abbr)

# show the data frame
data_brfss_smart2010
```


2. Answer the question: In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
more_than_7locations =
  data_brfss_smart2010 %>% 
  group_by(year, state) %>%
  distinct(location) %>% 
  count(name = "count") %>% 
  filter(year %in% c("2002", "2010") & count >= 7)

```
`r knitr::kable(more_than_7locations)`
In 2002, CT, FL, MA, NJ, NC, PA were observed at 7 or more locations.

In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations.



3. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state 

```{r, message=FALSE}
# filter the excellent rows and calculate the mean
excellent_responses =
  data_brfss_smart2010 %>% 
  filter(response == "Excellent") %>%
  select(year, state, response, data_value) %>%
  group_by(state, year) %>% 
  summarise(mean(data_value, na.rm = TRUE)) %>% # ignore the missing data when calculating
  rename("averages_data_value" = "mean(data_value, na.rm = TRUE)") %>% 
  arrange(year)


# make the plot
excellent_responses %>% 
  ggplot(aes(x = year, y = averages_data_value, group = state)) +
  geom_line(aes(color = state)) +
  labs(title = "spaghetti plot for average value over time within a state", 
       x = "year", 
       y = "average data value")
```

4. Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
NY_state_df = 
  data_brfss_smart2010 %>% 
  filter(state == "NY" & year %in% c(2006, 2010))

# make the density plot
NY_state_df %>% 
  group_by(response) %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .5, adjust = .5, color = "blue") +
  labs(
    title = "Two-Panel Plot for Responses in NY State in 2006 and 2010",
    x = "Data Value",
    y = "Density") +
  facet_grid( ~ year)
```

## Problem 3

1. Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r, message=FALSE}
accel_df = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(weekday_vs_weekend = day) %>% 
  mutate(weekday_vs_weekend = recode(weekday_vs_weekend,
    "Friday" = "Weekday",
    "Monday" = "Weekday",
    "Thursday" = "Weekday",
    "Tuesday" = "Weekday",
    "Wednesday" = "Weekday",
    "Saturday" = "Weekend",
    "Sunday" = "Weekend")) %>% 
  mutate( day = factor(day, 
    levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
  relocate(weekday_vs_weekend) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_prefix = "activity_",
    names_to = "minutes_order",
    values_to = "activity_counts"
  ) %>% 
  mutate(minutes_order = as.numeric(minutes_order)) %>% 
  relocate(week, day_id)
```



2.Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
total = 
  accel_df %>% 
  group_by(day_id) %>% 
  mutate(total_count = sum(activity_counts)) %>% 
  select(-activity_counts, -minutes_order) %>% 
  distinct() %>% 
  arrange(week, day)  

```
`r knitr::kable(total)`


```{r}
# to get the trends from the dataset
total %>% 
  ggplot(aes(x = day_id, y = total_count, group = weekday_vs_weekend)) +
  geom_point() +
  geom_line(aes(color = weekday_vs_weekend)) +
  labs(title = "plot for total activity counts in a week", 
       x = "day",
       y = "total count")
```

Trends: The total activity counts in every weekend tended to decrease when time went by.


3. Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
# mutate the variable to better fit in 24-hour range
accel_df %>% 
  mutate(hours_order = minutes_order/60) %>% 
  arrange(day_id) %>% 
# make a plot to show 24-hour activity time courses for each day
  ggplot(aes(x = hours_order, y = activity_counts, group = day_id)) +
  geom_line(aes(color = day), alpha = .5) +
  labs(title = "plot for 24-hour activity time courses for each day", 
       x = "24-hour of the day",
       y = "activity count") +
  scale_x_continuous(
    breaks = c(0, 4, 8, 12, 16, 20, 24),
    labels = c("0", "4", "8", "12", "16", "20", "24"))
```

Conclusion: the plot above showed that this man tended to have more activity counts around 12:00PM, between 8:00PM and 10:00PM in a day. Activity counts were in low status in the evening between 12:00AM and 4:00AM in a day, with a possible reason that he might be sleeping.



